Cell Decomposition of Rank 2 Quiver Grassmannians

### 2017-04-04 Grassmannian_cells.tex
Lemma 2.1, 2.7, Proposition 2.1
Added Lemma 3.4 which will provide the induction step for establishing cell decompositions, changed theorem numbering

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{definition}{Definition}[theorem]
\newtheorem{lemma}{Lemma}[theorem]
\newtheorem{question}{Question}[theorem]
\newtheorem{proposition}{Proposition}[theorem]

@@

\begin{lemma}
  For any $m\ge1$, we have 
  \[\dim\Hom(P_m,P_m)=1,\quad \dim\Hom(P_m,P_{m+1})=n,\quad \dim\Hom(P_{m+1},P_m)=0.\]  
  Moreover, for $1\le k<n$ and linearly independent functions $f_1,\ldots,f_k\in\Hom(P_m,P_{m+1})$, the map $[f_1\ \cdots\ f_k]:P_m^k\to P_{m+1}$ is injective. 
\end{lemma}
\begin{proof} Note that we have a Auslander-Reiten sequence
\[0\longrightarrow P_{m-1}\longrightarrow P_m^n\stackrel{[f_1\ \cdots\ f_n]}{\longrightarrow} P_{m+1}\longrightarrow 0\]
with irreducible homomorphisms $f_i$ which are also linear independent. Thus they do not factor through a representation $Z\neq P_{m}$ which already means that they are injective, see for instance \cite[Lemma 1.6]{ass}. If we pick $k<n$ linear independent homomorphisms $f_{i_1},\ldots, f_{i_k}$, this also means that $[f_{i_1}\ \cdots\ f_{i_k}]:P_m^k\to P_{m+1}$ is injective. Indeed otherwise, at least one of the homomorphisms were forced to factor through a representation $Z\neq P_{m}$.

 Write $P_{m+1}^{\{f_1,\ldots,f_k\}}$ for the cokernel of the map $[f_1\ \cdots\ f_k]:P_m^k\to P_{m+1}$, i.e. we have a short exact sequence
\begin{equation}
\label{eq:truncated preprojectives}
  0\longrightarrow P_m^k\stackrel{[f_1\ \cdots\ f_k]}{\longrightarrow} P_{m+1}\longrightarrow P_{m+1}^{\{f_1,\ldots,f_k\}}\longrightarrow 0.
\end{equation}
\end{proof}
\begin{lemma}
  For $m\ge1$ and linearly independent functions $f_1,\ldots,f_k\in\Hom(P_m,P_{m+1})$, the following hold:
  \begin{enumerate}
    \item $\Hom(P_m,P_{m+1}^{\{f_1,\ldots,f_k\}})$ is naturally isomorphic to $\Hom(P_m,P_{m+1})/\kk f_1\oplus\cdots\oplus\kk f_k$;
    \item $\Ext(P_{m+1}^{\{f_1,\ldots,f_k\}},P_m)$ is naturally isomorphic to $\kk f_1\oplus\cdots\oplus\kk f_k$;
    \item $\Hom(P_{m+1}^{\{f_1,\ldots,f_k\}},P_m)=0$.
  \end{enumerate}
\end{lemma}
\begin{proof}
  Applying the functor $\Hom(P_m,-)$ to the sequence \eqref{eq:truncated preprojectives} gives the exact sequence
  \[0\longrightarrow \Hom(P_m,P_m^k)\stackrel{[f_1\ \cdots\ f_k]\circ}{\longrightarrow} \Hom(P_m,P_{m+1})\longrightarrow \Hom(P_m,P_{m+1}^{\{f_1,\ldots,f_k\}})\longrightarrow 0,\]
  establishing (1).

  To see (2), we apply the functor $\Hom(-,P_m)$ to the sequence \eqref{eq:truncated preprojectives} to get an isomorphism $\Hom(P_m^k,P_m)\cong\Ext(P_{m+1}^{\{f_1,\ldots,f_k\}},P_m)$.
  Observe that $\Hom(P_m^k,P_m)$ is spanned by the coordinate projections $\pi_i:P_m^k\to P_m$ for $1\le i\le k$ and each of these gives rise to a commutative diagram
  \[\xymatrix{ & 0\ar[d] & 0\ar[d] & & \\ & P_m^{k-1}\ar[d]\ar@{=}[r]& P_m^{k-1}\ar[d]^{[f_1\ \cdots\ \widehat{f_i}\ \cdots\ f_k]}& & \\
    0\ar[r]& P_m^k\ar[r]^{[f_1\ \cdots\ f_k]}\ar[d]^{\pi_i}\ar@{}[dr]|{\ulcorner}& P_{m+1}\ar[r]\ar[d]& P_{m+1}^{\{f_1,\ldots,f_k\}}\ar[r]\ar@{=}[d]& 0 \\
    0\ar[r]& P_m\ar[r]^(.35){f_i}\ar[d]& P_{m+1}^{\{f_1,\ldots,\widehat{f_i},\ldots,f_k\}}\ar[r]\ar[d]& P_{m+1}^{\{f_1,\ldots,f_k\}}\ar[r]& 0\\
     & 0 & 0 & & }\]
  The bottom row of this diagram is the extension in $\Ext(P_{m+1}^{\{f_1,\ldots,f_k\}},P_m)$ naturally identified with the generator $f_i$ of $\kk f_1\oplus\cdots\oplus\kk f_k$.

@@

@@

\begin{lemma}
  \begin{enumerate}
    \item $\tilde P_m^{(n-1)}\cong\tilde P_{m-1}^{(1)}$
    \item The subrepresentation $\bigoplus_{i=1}^{k-1}\tilde P_{m-1,i}\oplus\bigoplus_{i=1}^k\tilde P_{m-2,i}\subset\bigoplus{i=1}^k\tilde P_{m-1,i}$ is in $(\tilde P_m^{(k)})^\perp$ and
      \[\Ext(\bigoplus_{i=1}^k\tilde P_{m-1,i},\tilde P_m^{(k)})\cong\Ext(\tilde P_{m-1}^{(k)},\tilde P_m^{(k)})\]
      where $\tilde P_{m-1}^{(k)}$ above denotes the cokernel of the inclusion.
  \end{enumerate}
\end{lemma}

\begin{lemma}\label{quotient}
 Consider $\ses{\tilde P_m}{\tilde P_{m+1}^{(k)}}{\tilde P_{m+1}^{(k-1)}}$. If $U\subset P_m$ such that $\Ext(\tilde P_{m+1}^{(k-1)},U)\neq 0$, then we have $\Ext(\tilde P_{m+1}^{(k-1)},\tilde P_m/U)= 0$.
\end{lemma}
\begin{proof}
First observe that $\tilde P_m\in ~^\perp \tilde P_{m+1}^{(k-1)}$ and $\Ext(\tilde P_{m+1}^{(k-1)},\tilde P_m)=k$. As $U$ is preprojective with $U\subset \tilde P_m\subset \tilde P_{m+1}^{(k)}$, it follows that $\Hom(U,\tilde P_{m+1}^{(k-1)})=0$ (check again). This yields $\Ext(\tilde P_m/U,\tilde P_{m+1}^{(k-1)})=0$. Consider
\[0\longrightarrow[\tilde P_{m+1}^{(k-1)},\tilde P_m/U]\longrightarrow[\tilde P_{m+1}^{(k-1)},U]^1\xlongrightarrow{\pi}[\tilde P_{m+1}^{(k-1)},\tilde P_m]^1\longrightarrow[\tilde P_{m+1}^{(k-1)},\tilde P_m/U]^1\longrightarrow 0.\]
Since $\Ext(\tilde P_{m+1}^{(k-1)},\tilde P_m)=k$, we have that $\pi$ is surjective if $\pi\neq 0$. If $U=U_1\oplus\ldots\oplus U_r$ is a decomposition into indecomposable representations, we can write $\pi=(\pi_1,\ldots,\pi_r)$ and it suffices to show that $\pi_j\neq 0$ for at least one $j=1,\ldots,r$. Thus we can without loss of generality assume that $U$ is indecomposable (check again). This means that $U\cong \tilde P_{i}$ for some $i=1,\ldots,m$. Moreover, $\tilde P_{m+1}/U$ is indecomposable (details).

@@

>>>>

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{question}[theorem]{Question}
\newtheorem{proposition}[theorem]{Proposition}

@@

\begin{lemma}
  For any $m\ge1$, we have 
  \[\dim\Hom(P_m,P_m)=1,\quad \dim\Hom(P_m,P_{m+1})=n,\quad \dim\Hom(P_{m+1},P_m)=0.\]  
  Moreover, for $1\le k<n$ and linearly independent functions $f_1,\ldots,f_k\in\Hom(P_m,P_{m+1})$, the map $[f_1\ \cdots\ f_k]:P_m^k\to P_{m+1}$ is injective. 
\end{lemma}
\begin{proof} 
  Note that we have an Auslander-Reiten sequence
  \begin{equation}
  \label{eq:AR sequence}
    0\longrightarrow P_{m-1}\longrightarrow P_m^n\stackrel{[f_1\ \cdots\ f_n]}{\longrightarrow} P_{m+1}\longrightarrow 0
  \end{equation}
  with irreducible homomorphisms $f_i$ which are also linearly independent. 
  Thus they do not factor through a representation $Z\neq P_{m}$ which already means that they are injective, see for instance \cite[Lemma 1.6]{ass}. 
  If we pick $k<n$ linear independent homomorphisms $f_{i_1},\ldots, f_{i_k}$, this also means that $[f_{i_1}\ \cdots\ f_{i_k}]:P_m^k\to P_{m+1}$ is injective. 
  Indeed otherwise, at least one of the homomorphisms were forced to factor through a representation $Z\neq P_{m}$.
\end{proof}

Write $P_{m+1}^{\{f_1,\ldots,f_k\}}$ for the cokernel of the map $[f_1\ \cdots\ f_k]:P_m^k\to P_{m+1}$, i.e. we have a short exact sequence
\begin{equation}
\label{eq:truncated preprojectives}
  0\longrightarrow P_m^k\stackrel{[f_1\ \cdots\ f_k]}{\longrightarrow} P_{m+1}\longrightarrow P_{m+1}^{\{f_1,\ldots,f_k\}}\longrightarrow 0.
\end{equation}

\begin{lemma}
  For $m\ge1$ and linearly independent functions $f_1,\ldots,f_k\in\Hom(P_m,P_{m+1})$, the following hold:
  \begin{enumerate}
    \item $\Hom(P_m,P_{m+1}^{\{f_1,\ldots,f_k\}})$ is naturally isomorphic to $\Hom(P_m,P_{m+1})/\kk f_1\oplus\cdots\oplus\kk f_k$;
    \item $\Ext(P_{m+1}^{\{f_1,\ldots,f_k\}},P_m)$ is naturally isomorphic to $\kk f_1\oplus\cdots\oplus\kk f_k$;
    \item $\Hom(P_{m+1}^{\{f_1,\ldots,f_k\}},P_m)=0$.
  \end{enumerate}
\end{lemma}
\begin{proof}
  Applying the functor $\Hom(P_m,-)$ to the sequence \eqref{eq:truncated preprojectives} gives the exact sequence
  \[0\longrightarrow \Hom(P_m,P_m^k)\stackrel{[f_1\ \cdots\ f_k]\circ}{\longrightarrow} \Hom(P_m,P_{m+1})\longrightarrow \Hom(P_m,P_{m+1}^{\{f_1,\ldots,f_k\}})\longrightarrow 0,\]
  establishing (1).

  To see (2), we apply the functor $\Hom(-,P_m)$ to the sequence \eqref{eq:truncated preprojectives} to get an isomorphism $\Hom(P_m^k,P_m)\cong\Ext(P_{m+1}^{\{f_1,\ldots,f_k\}},P_m)$.
  Observe that $\Hom(P_m^k,P_m)$ is spanned by the coordinate projections $\pi_i:P_m^k\to P_m$ for $1\le i\le k$ and each of these gives rise to a commutative diagram
  \[\xymatrix{ & 0\ar[d] & 0\ar[d] & & \\ & P_m^{k-1}\ar[d]\ar@{=}[r]& P_m^{k-1}\ar[d]^{[f_1\ \cdots\ \widehat{f_i}\ \cdots\ f_k]}& & \\
    0\ar[r]& P_m^k\ar[r]^{[f_1\ \cdots\ f_k]}\ar[d]^{\pi_i}\ar@{}[dr]|(.35){\ulcorner}& P_{m+1}\ar[r]\ar[d]& P_{m+1}^{\{f_1,\ldots,f_k\}}\ar[r]\ar@{=}[d]& 0 \\
    0\ar[r]& P_m\ar[r]^(.35){f_i}\ar[d]& P_{m+1}^{\{f_1,\ldots,\widehat{f_i},\ldots,f_k\}}\ar[r]\ar[d]& P_{m+1}^{\{f_1,\ldots,f_k\}}\ar[r]& 0\\
     & 0 & 0 & & }\]
  The bottom row of this diagram is the extension in $\Ext(P_{m+1}^{\{f_1,\ldots,f_k\}},P_m)$ naturally identified with the generator $f_i$ of $\kk f_1\oplus\cdots\oplus\kk f_k$.

@@

\begin{lemma}
  Let $f_1,\ldots,f_n\in\Hom(P_m,P_{m+1})$ be linearly independent morphisms and write $g:P_{m-1}\to P_m$ for the composition of the inclusion from \eqref{eq:AR sequence} with the projection $\pi_n:P_m^n\to P_m$ to the $n^{th}$ factor.  Then $P_{m+1}^{\{f_1,\ldots,f_{n-1}\}}\cong P_m^{\{g\}}$.
\end{lemma}
\begin{proof}
  We have the following commutative diagram
  \[\xymatrix{0\ar[r]& P_m^{n-1}\ar@{=}[d]\ar[r]& P_m^n\ar[r]^{\pi_n}\ar[d]^{[f_1\ \cdots\ f_n]}\ar@{}[dr]|(.65){\lrcorner}& P_m\ar[r]\ar[d]^{f_n}& 0 \\
  0\ar[r]& P_m^{n-1}\ar[r]^{[f_1\ \cdots\ f_{n-1}]}& P_{m+1}\ar[r]& P_{m+1}^{\{f_1,\ldots,f_{n-1}\}}\ar[r]& 0}\]
  Since the vertical morphism on the left is an equality, the right hand square is a pullback.  
  But this implies the kernels of the vertical morphisms coincide.  
  The middle vertical morphism fits into the exact sequence \eqref{eq:AR sequence} and so the kernel is $P_{m-1}$.
  It follows that there is a morphism $g:P_{m-1}\to P_m$ so that $P_{m+1}^{\{f_1,\ldots,f_{n-1}\}}\cong P_m^{\{g\}}$.
\end{proof}

@@

\begin{lemma}
  \begin{enumerate}
    \item $\tilde P_m^{(n-1)}\cong\tilde P_{m-1}^{(1)}$
    \item The subrepresentation $\bigoplus_{i=1}^{k-1}\tilde P_{m-1,i}\oplus\bigoplus_{i=1}^k\tilde P_{m-2,i}\subset\bigoplus_{i=1}^k\tilde P_{m-1,i}$ is in $(\tilde P_m^{(k)})^\perp$ and
      \[\Ext(\bigoplus_{i=1}^k\tilde P_{m-1,i},\tilde P_m^{(k)})\cong\Ext(\tilde P_{m-1}^{(k)},\tilde P_m^{(k)})\]
      where $\tilde P_{m-1}^{(k)}$ above denotes the cokernel of the inclusion.
  \end{enumerate}
\end{lemma}

\begin{lemma}\label{quotient}
 Consider $\ses{\tilde P_m}{\tilde P_{m+1}^{(k)}}{\tilde P_{m+1}^{(k-1)}}$. If $U\subset P_m$ such that $\Ext(\tilde P_{m+1}^{(k-1)},U)\neq 0$, then we have $\Ext(\tilde P_{m+1}^{(k-1)},\tilde P_m/U)= 0$.
\end{lemma}
\begin{proof}
First observe that $\tilde P_m\in ~^\perp \tilde P_{m+1}^{(k-1)}$ and $\Ext(\tilde P_{m+1}^{(k-1)},\tilde P_m)=k$. Assume that $U$ is preprojective with $\Hom(U,\tilde P_{m+1}^{(k-1)})=0$ (check again). This yields $\Ext(\tilde P_m/U,\tilde P_{m+1}^{(k-1)})=0$. Consider
\[0\longrightarrow[\tilde P_{m+1}^{(k-1)},\tilde P_m/U]\longrightarrow[\tilde P_{m+1}^{(k-1)},U]^1\xlongrightarrow{\pi}[\tilde P_{m+1}^{(k-1)},\tilde P_m]^1\longrightarrow[\tilde P_{m+1}^{(k-1)},\tilde P_m/U]^1\longrightarrow 0.\]
Since $\Ext(\tilde P_{m+1}^{(k-1)},\tilde P_m)=k$, we have that $\pi$ is surjective if $\pi\neq 0$. If $U=U_1\oplus\ldots\oplus U_r$ is a decomposition into indecomposable representations, we can write $\pi=(\pi_1,\ldots,\pi_r)$ and it suffices to show that $\pi_j\neq 0$ for at least one $j=1,\ldots,r$. Thus we can without loss of generality assume that $U$ is indecomposable (check again). This means that $U\cong \tilde P_{i}$ for some $i=1,\ldots,m$. Moreover, $\tilde P_{m+1}/U$ is indecomposable (details).

@@

Case $\Hom(U,P)\neq 0$?
